---
title: "UP data for meta-analysis stats"
author: "Raissa Mendonca"
date: "18 Apr 2018"
output: html_document
---

<span style="color:blue">*Last update on `r format(Sys.Date(), "%d %b %Y")`*</span>  

***
UP data using low-binding and high-binding sediments in Big Pup Creek and Salmon Trout to measure:
  
  1) Biofilm colozination on NDS cups amended with Cu and Ni at different concentrations. NPP, chl*a*, and metal bioaccumulation measured after  4 weeks
  2) Benthic invertebrate colonization on trays with the same metal amendments for 28 days. Inverts identified to family.
  3) Cotton strip decomposition

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
options(device=quartz)
#options(device="RStudioGD")
par(pin=c(2.5,2.5),las=1)
```

```{r Set WD & load libraries, include=F}
setwd("~/Documents/Kent State University/Costello Lab/GitHub/str-fun/UP_data")
library(vegan)
library(boot)
library(metafor)
library(lme4)
library(nlme)
```

```{r Graphing code, include=F}
logaxis = function(minlog,maxlog,side){
  pow <- seq(minlog,maxlog,by=1)
  ticksat <- as.vector(sapply(pow, function(p) (1:10)*10^p))
  axis(side, 10^pow, las=1, cex.axis=0.9)
  axis(side, ticksat, labels=NA, tcl=-0.25, lwd=0, lwd.ticks=1)
}

logaxis.b = function(minlog,maxlog,side){
  pow <- seq(minlog,maxlog,by=1)
  ticksat <- as.vector(sapply(pow, function(p) (1:10)*10^p))
  axis(side, 10^pow,labels=NA, las=1)
  axis(side, ticksat, labels=NA, tcl=-0.25, lwd=0, lwd.ticks=1)
}
```

```{r Load datasets, include=F}

biof <- read.csv("biofilm_data_elementa.csv")
    #Biofilm colonization on NDS cups, metal accum, chla and NPP
  biof$TOTMe <- (biof$Sed_Cu/63.54)+(biof$Sed_Ni/58.69)
    #Metal concentration variable as the sum of molar concentrations of both amended metals
  
  #Low-binding sediment
  biof.LBBP <- subset(biof,biof$Sed=="LB"&biof$Site=="BP")
  biof.LBST <- subset(biof,biof$Sed=="LB"&biof$Site=="ST")
  #High-binding sediment
  biof.HBBP <- subset(biof,biof$Sed=="HB"&biof$Site=="BP")
  biof.HBST <- subset(biof,biof$Sed=="HB"&biof$Site=="ST")


cott <- read.csv("cotton_data_elementa.csv")
    #Cotton strip decomposition data
  
  #Low-binding sediment
  cott.LBBP <- subset(cott,cott$Sed=="LB"&cott$Site=="BP")
  cott.LBST <- subset(cott,cott$Sed=="LB"&cott$Site=="ST")
  #High-binding sediment
  cott.HBBP <- subset(cott,cott$Sed=="HB"&cott$Site=="BP")
  cott.HBST <- subset(cott,cott$Sed=="HB"&cott$Site=="ST")
  
  
bugs <- read.csv("benthic_data_elementa.csv") 
    #Benthic invert family counts, total abund, richness, shannon div, inv simpson div and EPT abund
  bugs$TOTMe <- (bugs$Sed_Cu/63.54)+(bugs$Sed_Ni/58.69) 
    #Metal concentration variable as the sum of molar concentrations of both amended metals

bugs.hel <- cbind(bugs[,c(1:7,56)], decostand(bugs[,8:50],"hellinger",na.rm=T))
    #Hellinger transformation of invert count data to use in RDA
  bugs.hel <- na.omit(bugs.hel) #Remove NA

  #Low-binding sediment
  bugs.LBBP <- subset(bugs.hel,bugs.hel$Sed=="LB"&bugs.hel$Site=="Big Pup")
  bugs.LBST <- subset(bugs.hel,bugs.hel$Sed=="LB"&bugs.hel$Site=="Salmon-Trout")
  #High-binding sediment
  bugs.HBBP <- subset(bugs.hel,bugs.hel$Sed=="HB"&bugs.hel$Site=="Big Pup")
  bugs.HBST <- subset(bugs.hel,bugs.hel$Sed=="HB"&bugs.hel$Site=="Salmon-Trout")
  
mock <- read.csv("mock_mixed.csv")
  #Mock dataset to play with metafor package & functions // 1 fixed effect (habitat) and 1 random (study)
```

###1. Net primary production
```{r Linear models for LB & HB NPP, echo=F, tidy=TRUE, results="hide"}

  #Low-binding sediment

    ##Big Pup
npp.lbbp <- lm(NPP~log(TOTMe),data=biof.LBBP,na.action=na.omit)
  shapiro.test(resid(npp.lbbp)) #Ok
  summary(npp.lbbp) #No significant effect of metals (p=0.935, R2<0.01)
  
    ##Salmon Trout
npp.lbst <- lm(NPP~log(TOTMe),data=biof.LBST)
  shapiro.test(resid(npp.lbst)) #Ok
  summary(npp.lbst) #No significant effect of metals (p=0.739, R2<0.01)

  
  #High-binding sediment
    
    ##Big Pup
npp.hbbp <- lm(NPP~log(TOTMe),data=biof.HBBP)
  shapiro.test(resid(npp.hbbp)) #Ok
  summary(npp.hbbp) #No significant effect of metals (p=0.398, R2=0.03)

  
    ##Salmon Trout
npp.hbst <- lm(NPP~log(TOTMe),data=biof.HBST)
  shapiro.test(resid(npp.hbst)) #Ok
  summary(npp.hbst) #Significant effect of metals (p < 0.001, R2 = 0.60)

```

Log-linear models for NPP and sediment metal concentrations for low-binding and high-binding sediments

####Low-binding sediments
```{r Plots for LB NPP linear model, echo=F, tidy=TRUE, results="hide", fig.align='center'}
par(pin=c(2.5,2.5),las=1, cex.axis=0.9, cex.lab=0.9, cex.main=0.9)

plot(NPP~TOTMe,data=biof.LBBP,log="x",col="blue",xlim=c(1,10), xaxt="n", 
     xlab=expression(paste("Sediment metal (µmol g"^"-1"," dw)")), ylim=c(-0.1,0.4))
  points(NPP~TOTMe,data=biof.LBST,col="red", pch=6)
  logaxis(0,1,1)
  legend("topright",c("Big Pup Creek","Salmon Trout"), col=c("blue","red"), pch=c(1,6), cex=0.8)
  title("Low-binding sediment")

```

*NPP SLRs*

| Site | p-value | R<sup>2</sup> |
|:-:|:-:|:-:|
| Big Pup Creek | 0.935 | <0.01 | 
| Salmon Trout | 0.739 | <0.01 |

------

####High-binding sediments
```{r Plots for HB NPP linear models, echo=F, tidy=TRUE, results="hide", fig.align='center'}
par(pin=c(2.5,2.5),las=1, cex.axis=0.9, cex.lab=0.9, cex.main=0.9)

plot(NPP~TOTMe,data=biof.HBBP,log="x",col="blue",xlim=c(1,15), xaxt="n", 
     xlab=expression(paste("Sediment metal (µmol g"^"-1"," dw)")), ylim=c(-0.1,0.4))
  points(NPP~TOTMe,data=biof.HBST,col="red", pch=6)
  logaxis(0,2,1)
  legend("topright",c("Big Pup Creek","Salmon Trout"), col=c("blue","red"), pch=c(1,6), cex=0.8)
  title("High-binding sediment")
  abline(lm(NPP~log10(TOTMe),data=biof.HBST),col="red",lty=2)
  
```

*NPP SLRs*

| Site | p-value | R<sup>2</sup> |
|:-:|:-:|:-:|
| Big Pup Creek | 0.398 | 0.03 | 
| Salmon Trout | **<0.001** | 0.60 |

------


###2. Benthic invertebrate community composition
```{r RDA LB & HB models for benthic inverts, echo=F, tidy=TRUE, results="hide", fig.align='center'}

  #Low-binding sediment

    ##Big Pup
rda.lbbp <- rda(bugs.LBBP[,9:51]~log(TOTMe),bugs.LBBP)
  anova(rda.lbbp) #Model not significant (p=0.27)
  summary(rda.lbbp) #R2=0.11

    ##Salmon Trout
rda.lbst <- rda(bugs.LBST[,9:51]~log(TOTMe),bugs.LBST)
  anova(rda.lbst) #Model not significant (p=0.90)
  summary(rda.lbst) #R2=0.05


  #High-binding sediment
  
    ##Big Pup
rda.hbbp <- rda(bugs.HBBP[,9:51]~log(TOTMe),bugs.HBBP)
  anova(rda.hbbp) #Model not significant (p=0.26)
  summary(rda.hbbp) #R2=0.11

    ##Salmon Trout
rda.hbst <- rda(bugs.HBST[,9:51]~log(TOTMe),bugs.HBST)
  anova(rda.hbst) #Model not significant (p=0.50)
  summary(rda.hbst) #R2=0.10

```

RDA models for benthic invertebrate community and sediment metal concentrations for low-binding and high-binding sediments

####Low-binding sediments
```{r RDA plots for LB, echo=F, tidy=TRUE, results="hide", fig.align='center'}
par(pin=c(4,4),las=1, cex.axis=0.9, cex.lab=0.9, cex.main=0.9)

col.rda=c("sienna1","springgreen3","blue","goldenrod")[bugs.LBBP$Nom_Treat] 
    #Set colors based on metal treatments (ref, low, med, high)
pt.rda=c(16,8,23,2)[bugs.LBBP$Nom_Treat] 
    #Set points based on metal treatments (ref, low, med, high)


    ##Big Pup
plot(rda.lbbp,type="none",xlab="RDA1",ylab="PC1",xlim = c(-1.1,1.1), ylim=c(-1.1,1.1))
  title("Big Pup Creek")
  text(rda.lbbp,display="species",col=1,cex=0.9)
  points(rda.lbbp,pch=pt.rda,col=col.rda,cex=1.1,lwd=2) # Points instead of text for sites
  arrows(0,0,0.9,0, length=0.08, lwd=2)
  text(0.85,0.09,"log(total metals)",lwd=2,cex=0.75,font=2)
  legend("topright", c("Ref","Low","Med","High"), col=c("sienna1","springgreen3","blue","goldenrod"),
         pch=c(16,8,23,2))
  
    ##Salmon Trout
plot(rda.lbst,type="none",xlab="RDA1",ylab="PC1",xlim = c(-1.15,1.1), ylim=c(-1.1,1.1))
  title("Salmon Trout")
  text(rda.lbst,display="species",col=1,cex=0.9)
  points(rda.lbst,pch=pt.rda,col=col.rda,cex=1.1,lwd=2) # Points instead of text for sites
  arrows(0,0,0.8,0, length=0.08, lwd=2)
  text(0.88,0.09,"log(total metals)",lwd=2,cex=0.75,font=2)  
  legend("topright", c("Ref","Low","Med","High"), col=c("sienna1","springgreen3","blue","goldenrod"),
         pch=c(16,8,23,2))
```

*Benthic invert RDAs*

| Site | p-value | R<sup>2</sup> |
|:-:|:-:|:-:|
| Big Pup Creek | 0.27 | 0.11 | 
| Salmon Trout | 0.90 | 0.05 |

------

####High-binding sediments
```{r RDA plots for HB, echo=F, tidy=TRUE, results="hide", fig.align='center'}
par(pin=c(4,4),las=1, cex.axis=0.9, cex.lab=0.9, cex.main=0.9)

col.rda=c("sienna1","springgreen3","blue","goldenrod")[bugs.LBBP$Nom_Treat] 
    #Set colors based on metal treatments (ref, low, med, high)
pt.rda=c(16,8,23,2)[bugs.LBBP$Nom_Treat] 
    #Set points based on metal treatments (ref, low, med, high)
  

    ##Big Pup
plot(rda.hbbp,type="none",xlab="RDA1",ylab="PC1",xlim = c(-1.1,1.1), ylim=c(-1.1,1.1))
  title("Big Pup Creek")
  text(rda.hbbp,display="species",col=1,cex=0.9)
  points(rda.hbbp,pch=pt.rda,col=col.rda,cex=1.1,lwd=2) # Points instead of text for sites
  arrows(0,0,0.8,0, length=0.08, lwd=2)
  text(0.85,0.09,"log(total metals)",lwd=2,cex=0.75,font=2)
  legend("topright", c("Ref","Low","Med","High"), col=c("sienna1","springgreen3","blue","goldenrod"),
         pch=c(16,8,23,2))
  
  ##Salmon Trout
plot(rda.hbst,type="none",xlab="RDA1",ylab="PC1",xlim = c(-1.1,1.1), ylim=c(-1.1,1.1))
  title("Salmon Trout")
  text(rda.hbst,display="species",col=1,cex=0.9)
  points(rda.hbst,pch=pt.rda,col=col.rda,cex=1.1,lwd=2) # Points instead of text for sites
  arrows(0,0,0.9,0, length=0.08, lwd=2)
  text(0.85,0.09,"log(total metals)",lwd=2,cex=0.75,font=2)
  legend("topright", c("Ref","Low","Med","High"), col=c("sienna1","springgreen3","blue","goldenrod"),
         pch=c(16,8,23,2))
  
```

*Benthic invert RDAs*

| Site | p-value | R<sup>2</sup> |
|:-:|:-:|:-:|
| Big Pup Creek | 0.26 | 0.11 | 
| Salmon Trout |0.50 | 0.10 |

------


###3. Calculating standard error of R<sup>2</sup> based on solution found on [Stats Exchange](https://stats.stackexchange.com/questions/175026/formula-for-95-confidence-interval-for-r2)
```{r Rsqr std error function & lm bootstrap, include=F}
  #Function to calculate standard error of Rsqr
SER2 <- function(model){
      R2 <- summary(model)$r.squared
      k <- summary(model)$df[1]
      n <- summary(model)$df[1]+summary(model)$df[2]
      se <- sqrt(4*R2*(1-R2)^2*(n-k-1)^2/((n^2-1)*(n+3)))
      R2.ci <- round(data.frame(R2-(abs(qt(0.05/2, n-2))*se),
                                R2,
                                R2+(abs(qt(0.05/2, n-2))*se),
                                se),
                     digits=3)
      colnames(R2.ci) <- c("Lower","R2","Upper","SE")
      return(R2.ci)
  }

SER2(npp.lbbp)

  #Bootstrapped R2 and confidence interval from regression
boot.R2 <- boot(biof.LBBP,function(data,indices)
              summary(lm(NPP~log(TOTMe),data[indices,]))$r.squared, R=10000)
  boot.R2
  boot.R2$t0
  quantile(boot.R2$t,c(0.025,0.975))
  
```

```{r Rsqr std error function & RDA bootstrap, include=F}
  #Function to calculate standard error of Rsqr
SER2.rda <- function(model){
      RDA.R2 <- RsquareAdj(model)$r.squared
      amod <- anova(model)
      n <- amod[1]$Df[1]+amod[1]$Df[2]
      k <- amod[1]$Df[1]
      se <- sqrt((4*RDA.R2*((1-RDA.R2)^2)*((n-k-1)^2))/((n^2-1)*(n+3)))
      RDA.R2.ci <- round(data.frame(RDA.R2-(abs(qt(0.05/2, n-2))*se),
                                RDA.R2,
                                RDA.R2+(abs(qt(0.05/2, n-2))*se),
                                se),
                     digits=3)
      colnames(RDA.R2.ci) <- c("Lower","R2","Upper","SE")
      return(RDA.R2.ci)
  }

SER2.rda(rda.hbbp)
  
  #Bootstrapped R2 and confidence interval from RDA
boot.R2 <- boot(bugs.LBBP,function(data,indices)
              RsquareAdj(rda(bugs.LBBP[,9:51]~log(TOTMe),data[indices,]))$r.squared, 
              R=10000)
  boot.R2
  boot.R2$t0
  quantile(boot.R2$t,c(0.025,0.975))
  sd(boot.R2$t)

```

*NPP models*

|Treatment| R<sup>2</sup> | SE<sub>FORMULA</sub> | R<sup>2</sup> *boot* | SE<sub>BOOT</sub> |
|:-:|:-:|:-:|:-:|:-:|
| Low-binding Big Pup | 0.0004 | 0.007 | 0.0004 | 0.092 |
| Low-binding Salmon Trout | 0.005 | 0.024 | 0.005 | 0.067 |
| High-binding Big Pup | 0.033 | 0.059 | 0.033 | 0.075 |
| High-binding Salmon Trout | 0.602 | 0.104 | 0.602 | 0.125 |

--------

*RDA models*

|Treatment| R<sup>2</sup> | SE<sub>FORMULA</sub> | R<sup>2</sup> *boot* | SE<sub>BOOT</sub> |
|:-:|:-:|:-:|:-:|:-:|
| Low-binding Big Pup | 0.107 | 0.128 | 0.107 | 0.030 | 
| Low-binding Salmon Trout | 0.048 | 0.091 | 0.048 | 0.043 | 
| High-binding Big Pup | 0.106 | 0.128 | 0.106 | 0.036 |
| High-binding Salmon Trout | 0.097 | 0.125 | 0.097 | 0.032 |

--------

###4. Testing R<sup>2</sup> effect size using *metafor*
```{r Matrix with R2 variables for metafor, echo=F, tidy=T}

  #Models for FUN (lm) and STR (RDA)
FUNmod <- list(npp.lbbp,npp.lbst,npp.hbbp,npp.hbst)
STRmod <- list(rda.lbbp,rda.lbst,rda.hbbp,rda.hbst)
  #Datasets for FUN and STR
FUNdat <- list(biof.LBBP,biof.LBST,biof.HBBP,biof.HBST) 
STRdat <- list(bugs.LBBP,bugs.LBST,bugs.HBBP,bugs.HBST)

  #Matrix with FUN R2, FUN SE formula, FUN SE bootstrap, STR R2, STR SE formula, STR SE bootstrap
R2.sum <- data.frame(matrix(NA, nrow=4, ncol=12, 
                     dimnames=list(c("Low-binding Big Pup","Low-binding Salmon Trout",
                                     "High binding Big Pup","High-binding Salmon Trout"),
                                   c("FUN R2","FUN SE form","FUN SE boot",
                                     "STR R2","STR SE form","STR SE boot","R2 dif",
                                     "SE form pooled","SE boot pooled","Stream","Sed","Treat"))))

  R2.sum[1:4,1] <- sapply(FUNmod, function(x) RsquareAdj(x)$r.squared) #FUN R2
  R2.sum[1:4,2] <- sapply(FUNmod, function(x) SER2(x)$SE) #FUN SE formula
  R2.sum[1:4,3] <- sapply(FUNdat, function(x) sd(boot(x,function(data,indices)
                        summary(lm(NPP~log(TOTMe),data[indices,],na.action=na.omit))$r.squared, R=10)$t)) #FUN SE bootstrap
  R2.sum[1:4,4] <- sapply(STRmod, function(x) RsquareAdj(x)$r.squared) #STR R2
  R2.sum[1:4,5] <- sapply(STRmod, function(x) SER2.rda(x)$SE) #STR SE formula
  R2.sum[1:4,6] <- sapply(STRdat, function(x) sd(boot(x,function(data,indices)
                        RsquareAdj(rda(x[,9:51]~log(TOTMe),data[indices,]))$r.squared, R=10)$t)) #STR SE bootstrap
  R2.sum[1:4,7] <- R2.sum$FUN.R2-R2.sum$STR.R2 #R2 diff as FUN-STR (i.e., if negative, FUN smaller than STR)
  R2.sum[1:4,8] <- sqrt(R2.sum$FUN.SE.form^2+R2.sum$STR.SE.form^2) #Pooled SE formula as sqrt(SE1^2+SE2^2)
  R2.sum[1:4,9] <- sqrt(R2.sum$FUN.SE.boot^2+R2.sum$STR.SE.boot^2) #Pooled SE bootstrap
  R2.sum[1:4,10] <- as.character(c("Big Pup","Salmon Trout","Big Pup","Salmon Trout"))
  R2.sum[1:4,11] <- as.character(c("Low-bind","Low-bind","High-bind","High-bind"))
  R2.sum[1:4,12] <- as.character(c("LBBP","LBST","HBBP","HBST"))
  
R2.sum <- cbind(round(R2.sum[,1:9],4),R2.sum[,10:12])
R2.sum

#Dave is the best advisor == T
```

```{r Metafor analysis of R2 dif and pooled SE, echo=F, tidy=T}
  #Function to return selected summaries for an rma model
rma.sum <- function(model) {
            sumList <- list(summary.rma(model),
                            round(coef.summary.rma(summary.rma(model)),4),
                            anova.rma(model),
                            round(weights(model),3))
            names(sumList) <- c("summary.rma","coef.summary","anova.rma","weights")
            print(sumList)
}

  #Function 'rma.uni' used to fit fixed- and mixed-effects models via linear models in metafor
    ##Random effects in 'rma.uni' refers to the variance (tau^2) of the true effects/outcomes in a population of studies sampled from a larger collection of studies, assuming the population is a random sample of the larger collection
    ##In mixed-effects (moderators included), tau^2 denotes teh amount of residual heterogeneity in the true effects/outcomes, which is the amount of variation not accounted for by the moderators included in the model

fixedboot <- rma.uni(yi=R2.dif, sei=SE.boot.pooled, data=R2.sum, method="FE")
  rma.sum(fixedboot)
randomboot <- rma.uni(yi=R2.dif, sei=SE.boot.pooled, data=R2.sum, method="REML")    
  rma.sum(randomboot)
mixedboot <- rma.uni(yi=R2.dif~Stream, sei=SE.boot.pooled, data=R2.sum, method="REML")
  rma.sum(mixedboot)
  ##yi can be written as formula (yi ~ mod1 + mod2 + mod3), which will override any 'mods'

```

```{r Test with mock dataset in metafor, echo=T, tidy=T}
mock.uni <- rma.uni(yi = R2dif~Fixed, sei = R2dif_SE, data = mock)
  rma.sum(mock.uni)

mock.mv <- rma.mv(R2dif~Fixed, R2dif_SE^2, random= ~1|Study, data=mock)
  rma.sum(mock.mv)
```

####*Calculating weighted linear regressions with lme() based on solution found on* [Stats Exchange](https://stats.stackexchange.com/questions/70425/weights-argument-in-lm-and-lme-very-different-in-r-am-i-using-them-correctly)
```{r Test of weighted linear regression with lme4, echo=T, tidy=T}
mock.lme <- lme(R2dif~Fixed, random= ~1|Study, data=mock, weights= ~I(1/R2dif_SE))
mock.lme
```